---
title: "exploratory-data-analysis"
author: "Christian Burkhart"
output: html_document
---

# Exploratory Data Analysis

```{r, echo=FALSE}
options(warn=-1)
```


```{r Load libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(gridExtra)
library(scales)
library(psych)
# library(nFactors)
# library(FactoMineR)
```

```{r Colors, echo=FALSE}
color.one <- "#21313E"
color.two <- "#4CA173"
```

```{r Helper functions, echo=FALSE}
density.plot <- function(measures) {
  # Get columns of text measure variables
  col.nums <- match(measures, names(data))
  
  # Melt data frame to long format
  text.measures.long <- melt(dplyr::select(data, col.nums))
  
  # Assign new variable to data frame
  # Regex check if variable is pre or post
  text.measures.long$time <- as.factor(ifelse(grepl("pre", 
                                          text.measures.long$variable, 
                                          ignore.case = TRUE), "pre", "post"))
  text.measures.long$text <- as.factor(ifelse(grepl("diagnostic", 
                                          text.measures.long$variable, 
                                          ignore.case = TRUE), "diagnostic", "explanation"))
  
  # Plot coherence sentences
  ggplot(text.measures.long, aes(x = value, fill = time)) +
    geom_density(alpha = .4) +
    facet_grid( ~ text)
}
```


```{r Load data set}
data <- read.csv("../data/data-lecture-study.csv", sep = ";")
```

`r nrow(data)` subjects were included in the data analysis. Their mean age was `r round(mean(data$age), 2)` (*SD* = `r round(sd(data$age), 2)`). `r percent(round(as.numeric(table(data$gender)[2]) / nrow(data), 2))` were female. 

## Univariate Analysis

```{r Grid of important subjects variables, echo=FALSE}
histogram.age <- qplot(age, data = data, geom = "histogram", binwidth = 1)
histogram.university.semester <- qplot(university.semester, data = data, 
                                       geom = "histogram", binwidth = 2)
histogram.subject.semester <- qplot(subject.semester, data = data, 
                                    geom = "histogram", binwidth = 2)
histogram.preknowledge.teaching <- qplot(preknowledge.teaching, data = data, 
                                         geom = "histogram", binwidth = 1)
histogram.teaching.experience <- qplot(teaching.experience, data = data, 
                                       geom = "histogram", binwidth = 1)
histogram.scientific.writing <- qplot(scientific.writing, data = data, 
                                       geom = "histogram", binwidth = 1)
histogram.internship <- qplot(internship, data = data, 
                                       geom = "histogram", binwidth = 1)

grid.arrange(histogram.age, histogram.university.semester,
            histogram.subject.semester, histogram.preknowledge.teaching,
            histogram.teaching.experience, histogram.scientific.writing,
            histogram.internship, ncol = 2)
```

Most students are in their 20s with very few above or below. Some students used negative values in the variable university semester. We need to turn them to absolute values:

```{r Turn university.semester to absolute values}
data$university.semester <- abs(data$university.semester)
```

The students do not have a lot of teaching experience either in general knowledge about teaching or real teaching experience. Most students have never attended a course on scientific writing. As students are early in their studies most of them have not finished their internship as part of their teacher education program.

### Factor analysis coherence questionnaire

Our subjects filled out a self-made questionnaire on text coherence. We assumed two factors: local and global coherence. In order to test this hypothesis, we need to calculate a exploratory factor analysis:

```{r Factor analysis coherence questionnaire, echo=FALSE}
# Select all variables of text coherence questionnaire
coherence.data <- select(data, c(13, 14, 15, 16, 17, 18, 20, 22, 23, 24, 25))

# fit <- factanal(coherence.data, 2, rotation = "varimax")
# print(fit, digits = 2, cutoff = .3, sort = TRUE)
# load <- fit$loadings[, 1:2]
# plot(load, type = "n")
# text(load, labels = names(coherence.data), cex = .7)

# ev <- eigen(cor(coherence.data)) # get eigenvalues
# ap <- parallel(subject=nrow(coherence.data),var=ncol(coherence.data),
#   rep=100,cent=.05)
# nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
# plotnScree(nS)
# 
# result <- PCA(coherence.data)
```


### Univariate Data of texts

Let's see how the number of sentences differed for each text and variable.

```{r Histogram text data, echo=FALSE, message=FALSE}
# Save variable names in vector
sentences.variables <- 
  c("explanation.pre.num.sentences", "explanation.post.num.sentences",
    "diagnostic.pre.num.sentences", "diagnostic.post.num.sentences",
    "schoolsystems.num.sentences")

# Find columns that match variable names
col.nums <- match(sentences.variables, names(data))

# Select columns from data frame and bring data to long format
sentences <- melt(dplyr::select(data, col.nums))

# Draw density histogram of sentence variables
ggplot(sentences, aes(x = value, fill = variable)) +
  geom_density(alpha = .2)
```

Most students wrote about 10 to 15 sentences for each text. The least sentences were written in the last text on schoolsystems. Overall, there is not much difference except for the last text. 

Let's see how subjects different in the other text measures.

```{r Text measures, echo=FALSE}
# Save variable names in vectors
coherence.measures <- c("explanation.pre.coh.sen", "explanation.post.coh.sen",
                   "diagnostic.pre.coh.sen", "diagnostic.post.coh.sen")
non.coherence.measures <- c("explanation.pre.not.coh.sen", 
                            "explanation.post.not.coh.sen",
                            "diagnostic.pre.not.coh.sen",
                            "diagnostic.post.not.coh.sen")
cluster.measures <- c("explanation.pre.num.clusters",
                      "explanation.post.num.clusters",
                      "diagnostic.pre.num.clusters",
                      "diagnostic.post.num.clusters")
```

```{r Coherent sentences distribution, echo=FALSE, message=FALSE}
density.plot(coherence.measures)
```

All in all students seem to have written a slight amount of more coherent sentences in the first text on exlanation. The difference between their draft (pre) and their revision (post) is minor. There is a small trend towards more coherent sentences in the revision.

```{r Non-coherent sentences distribution, echo=FALSE, message=FALSE}
density.plot(non.coherence.measures)
```

Both distributions seem to be quite identical indicating that students non-coherent sentences didn't change with their revisions.

```{r Cluster measures, echo=FALSE, message=FALSE}
density.plot(cluster.measures)
```

Subjects had less clusters in their revision compared to their draft in both texts. 

```{r Duplicates, echo=FALSE}
explanation.duplicate.prop <- prop.table(table(data$explanation.duplicate))
diagnostic.duplicate.prop <- prop.table(table(data$diagnostic.duplicate))
```

`r percent(round(explanation.duplicate.prop[2], 4))` of subjects did not revise their texts at all in the first text on instructional explanations. `r percent(round(diagnostic.duplicate.prop[2], 4))` did not revise their text in the second text on teacher diagnostic. One reason is that students that wrote text with only a single calculated cluster received a feedback indicating that their texts are coherent. Subjects might be prompted to think that a revision of their drafts was not necessary under these circumstances. 

```{r Density plot global coherence}
global.coherence.variables <- c("explanation.global.coherence.pre",
                                            "explanation.global.coherence.post",
                                            "diagnostic.global.coherence.post",
                                            "diagnostic.global.coherence.pre")
density.plot(global.coherence.variables)
```

It seems as if global coherence improved a bit with the text's revision. 

```{r}
# Melt data frame to long format
col.nums <- match(global.coherence.variables, names(data))
global.coherence <- melt(dplyr::select(data, col.nums))

# Assign new variable to data frame
# Regex check if variable is pre or post
global.coherence$time <- as.factor(ifelse(grepl("pre", 
                                        global.coherence$variable, 
                                        ignore.case = TRUE), "pre", "post"))
global.coherence$text <- as.factor(ifelse(grepl("diagnostic", 
                                        global.coherence$variable, 
                                        ignore.case = TRUE), "diagnostic", "explanation"))

# Select variables
global.coherence <- select(global.coherence, value, time, text)
  
```


```{r Density plot local coherence}
local.cohesion.variables <- c("explanation.local.cohesion.pre",
                                            "explanation.local.cohesion.post",
                                            "diagnostic.local.cohesion.post",
                                            "diagnostic.local.cohesion.pre")
density.plot(local.cohesion.variables)
```

We calculated cohesion by the ratio between the number of clusters and the number of sentences for each text. The lower our measure of local coherence, the more locally cohesive the text. It seems as if local cohesion markedly improved in both texts. 



## Bivariate Analysis

## Multivariate Analysis



